--- 
title: "Economic Data Visualization"
author: "Dr. Lin, Mau-Ting"
date: "2021"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
---


```{r, eval=T, echo=F}
klippy::klippy()
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=F)
```


# Prerequisites

## Packages

```{r}
install.packages(
  c("showtext","extrafont", "remotes")
)
```

  * **showtext** and **extrafont** are for Chinese characters to show properly in the graph.

```{r}
remotes::install_github("tpemartin/econDV2", force=T)
```

## Start a project

You should always organize documents with the same purpose inside a unified folder, which is called a **project folder**, and initiate an **R project** with it.

Start a new project for your ggplot environment:

**Project** > **New Project...** > **New Directory** > **New Project**

<div class="alert alert-info">
Initiate a project and setup environment:

  * Addins > (under `econDV2`) Setup environment

</div>

## .Rprofile

```{r}
sysfonts::font_families()
sysfonts::font_add_google('Noto Sans TC')
sysfonts::font_families()
```

```{r}
library(ggplot2)
library(showtext)
library(econDV2)
sysfonts::font_add_google('Noto Sans TC')
showtext_auto()
theme_set(
  theme(
    text=element_text(family = "Noto Sans TC")
  )+
    theme_classic()
)

gg <- list(
  dash = econDV2::ggdash,
  geom = econDV2::ggbrowse,
  aes = econDV2::ggaes
)

```

## 其他{#otherSoftware}

<div class="alert alert-info">
  
  * 後製軟體： Choose your own image editor, but make sure it support SVG editing. Here I use [Boxy SVG](https://boxy-svg.com/).
  
  * 程式排版：
```{r}
install.packages("styler")
```
  

</div>

